---
title: "Markdown Assignment"
author: "Alex Rudenko"
date: "November 4, 2016"
output: html_document
---
Activity Monitoring Analysis
============================

## 1. Loading and preprocessing the data


```{r setup, include=TRUE,results='markup'}
knitr::opts_chunk$set(echo = TRUE)
raw_activity <- read.csv("activity.csv")
all_activity <- raw_activity[!is.na(raw_activity$steps),]
activity <- all_activity[all_activity$steps>0,]
str(activity)
#agg <- tapply(activity$steps,activity$date,sum)


```

## 2. Total number of steps taken per day, their mean and median
```{r}
# steps_agg <- tapply(activity$steps,activity$date,sum)
daily_steps <- aggregate(steps ~ date, activity, sum)
hist(daily_steps$steps, main="Histogram of the total number of steps taken each day")
steps_mean <- mean(daily_steps$steps, na.rm = TRUE)
steps_median <- median(daily_steps$steps, na.rm = TRUE)

```
Mean of number of steps taken each day is `r  steps_mean`

Median number of steps taken each day is `r steps_median`

## 3. Average number of steps taken

```{r}
library(ggplot2)
steps_avg_per_interval <- aggregate(steps ~ interval, activity,mean)
plot(steps_avg_per_interval$interval,steps_avg_per_interval$steps, type="l",xlab = "5-min intervals", ylab = "Average number of steps")
ms <- max(steps_avg_per_interval$steps)
x <- which.max(steps_avg_per_interval$steps)
intN <- steps_avg_per_interval$interval[x]
```

Interval with maximum average number of steps (`r ms`) is `r intN`

## 4. Imputing missing values

```{r}
tot_NA <- sum(!complete.cases(raw_activity))
```
Total number of missing values is `r tot_NA`

### Replacing missing steps values with average for that interval
```{r}
imputed <- raw_activity
avg <- aggregate(steps ~ interval, all_activity,mean) 
for(i in 1:length(imputed$steps)) {
  if (is.na(imputed$steps[i])) {
    x <- which(avg$interval == imputed$interval[i])
    imputed$steps[i] <- avg$steps[x]
  } 
}
imp_NA <- sum(!complete.cases(imputed))
```
Total number of missing values is `r imp_NA`

```{r}
daily_steps <- aggregate(steps ~ date, imputed, sum)
hist(daily_steps$steps, main="Total number of steps taken each day after imputing mising values")
new_mean <- mean(daily_steps$steps, na.rm = TRUE)
new_median <- median(daily_steps$steps, na.rm = TRUE)
d_mean <- new_mean - steps_mean
d_median <- new_median - steps_median

```
After the missing values where imputing:

Mean of number of steps taken each day: `r  new_mean` (changed by `r  d_mean`)

Median number of steps taken each day: `r new_median` (change by `r d_median`)

## Differences in activity patterns between weekdays and weekends

```{r}
dtw <- as.POSIXlt(as.Date(imputed$date,"%m/%d/%Y"))$wday
dt_type <- ifelse(dtw %in% c(1:5), "weekday", "weekend")
imputed$daytype <- factor(dt_type, labels = c("weekday", "weekend"))

# Creating a plot
library(lattice)
avg <- aggregate(steps ~ interval + daytype, imputed, mean)
y <- avg$steps
x <- avg$interval
f <- avg$daytype
xyplot(y ~ x | f, layout= c(1,2), type="l", xlab = "5-min intervals", ylab="Average number od steps")
```

## END

